---
title: "General Social Survey"
author: "Sandra Rogers"
date: "10/07/2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is no relationship between gun ownership and whether one is in favor of capital punishment.

$H_A$ : There is a relationship between gun ownership and whether one is in favor of capital punishment.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+OWNGUN, data=mydata4, main='Capital Punishment & Gun Ownership' , type='percent', xlab='Capital Punishment')

```

This barchart shows that of the people who favor capital punishment, gun ownership is about the same amount, whereas of the people that do not favor capital punishment, there are more people who do not own guns.



#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~CAPPUN+OWNGUN, data=mydata4)
rowPerc(tab1)
```
This information shows that of the people who favor capital punishment, gun ownership is about the same amount, whereas of the people that do not favor capital punishment, there are more people who do not own guns.


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisqtestGC(tab1)
```


The P-value for the chi-square test is 0.02 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the contention that gun ownership and whether or not it favors capital punishment are related.

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is no relationship between the big bang theory and whether one believes in the afterlife.

$H_A$ : There is a relationship between the big bang theory and whether one believes in the afterlife.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results
#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, main='The Big Bang Theory & The Afterlife', type='percent', xlab='Big bang theory')
```

This barchart shows that of the people who believe in the big bang theory many believe in the afterlife, similarly of the people that do not believe in the big bang theory, many believe in the afterlife.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(tab1)
```

This information shows that of the people who believe in the big bang theory many believe in the afterlife, similarly of the people that do not believe in the big bang theory, many believe in the afterlife.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```
The P-value for the chi-square test is 0.08 > 0.05 so we failed to reject the null hypothesis and conclude we have no evidence to support the contention that the belief in the big bang theory is related to the belief of the afterlife.


# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables in this analysis are categorical; hence we use bar graphs and tables to summarize the data, and a chi-squared test for inference.

$H_0$ : There is no relationship between the death penalty and ones race.

$H_A$ : There is a relationship between the the death penalty and ones race.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, main='Capital Punishment & Race', type='percent', xlab='Cappun')
```
This bar graph shows that a persons opinion on capital punishment is dependent on race. With whites and others, you can see that there is slightly more who favor capital punishment whereas with blacks, a lot more oppose capital punishment.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1 = xtabs(~CAPPUN+RACE, data=mydata4)
rowPerc(tab1)
```
This information shows that a persons opinion on capital punishment is dependent on race. With whites and others, you can see that there is slightly more who favor capital punishment whereas with blacks, a lot more oppose capital punishment.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab1)
```
The P-value for the chi-square test is 9e-04 < 0.05 so we reject the null hypothesis and conclude we have evidence to support the contention that people's opinions on capital punishment and whether race makes a difference is dependent.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Question One: After reviewing the bargraph and chi-square test we can say there is a relationship between gun ownership and whether one is in favor of capital punishment.

Question Two: After reviewing the bargraph and chi-square test we can say there is no relationship between the big bang theory and whether one believes in the afterlife.

Question Three: After reviewing the bargraph and chi-square test we can say there is a relationship between the death penalty and one’s race. 


